---
title: "State PCE (SAPCE 3) vs PRO Summary Use table F01 column comparison"
author: "JV"
date: "2025-12-17"
output: html_document
---

#1. Get State PCE data (SAPCE 3)

```{r}

rm(list = ls()) # Clear environment
devtools::load_all(".") 
devtools::load_all("../useeior/") 

stateior_path <- getwd()
script_path <- "data-raw"
SAPCE_level <- 3
year <- 2017

# Load State PCE from source data file and save to /data/ folder
source(file.path(stateior_path, script_path, "BEAData.R"))

# Read PCE data from /data/ folder and confirm we are using SAPCE 3 data (more detailed line items)
State_PCE <- readRDS("data/State_PCE_2017_0.4.0.rds")
if(length(unique(State_PCE$LineCode)) ==  24){
  print("Using SAPCE 1")
}else if(length(unique(State_PCE$LineCode)) == 113){
  print("Using SAPCE 3")
}else{
  print("Using unknown SAPCE level")
}

US_PCE <- State_PCE[State_PCE$GeoName=="United States",]

```


#2. Get Use table before redefinitions as used in the state models
```{r}

# Define relevant spec
specs <- list()
specs$BaseIOSchema <- 2017
specs$model_ver <- '0.4.0' 

# Get relevant national Use table
U <- getNationalUse(iolevel = "Summary", year = year, specs = specs)

# Isolate final demand column (PCE) as its own df
U_F10 <- U[,colnames(U)=="F010", drop = FALSE]
U_F10 <- cbind(rownames(U_F10), U_F10)
colnames(U_F10) <- c("CommodityCode","U_Prod_F10")
U_F10$U_Prod_F10 <- U_F10$U_Prod_F10/1e6 # to align with unit from pce brigde


```


#3. Get PCE Summary Bridge File and sort by commodity and producer value to compare with Use table F10 values
```{r}

year <- 2017 

# Download BEA Summary PCE bridge table
bridge_file <- "inst/extdata/PCEBridge_Summary.xlsx"
if(!file.exists(bridge_file)) {
  download.file("https://apps.bea.gov/industry/release/xlsx/PCEBridge_Summary.xlsx",
                bridge_file, mode = "wb")
}
# Note that the first column is actually called NIPA Line but it is called here LineCode to match with 
# the State_PCE column name

column_names <- c("LineCode", "PCECategory", "CommodityCode", "CommodityDescription",
                  "ProducersValue", "Transportation Costs", "Wholesale", "Retail", "PurchasersValue","Year")

PCEBridge_Summary <- read_excel(bridge_file, as.character(year))

# Remove the first 4 columns as they do not contain relevant data
PCEBridge_Summary <- PCEBridge_Summary[-(1:4),]
colnames(PCEBridge_Summary) <- column_names

# Sort and find ratios by Commodity Code
PCEBridge_Summary_byCom <- PCEBridge_Summary[order(PCEBridge_Summary$CommodityCode),]

# Find sum of ProducersValue and PurchasersValue by Commodity
# First convert to numeric
PCEBridge_Summary_byCom$ProducersValue <- as.numeric(PCEBridge_Summary_byCom$ProducersValue)
PCEBridge_Summary_byCom[which(is.na(PCEBridge_Summary_byCom$ProducersValue))] <- 0 # replace possbile NAs from character conversion with 0s

# Find producer and purchaser sums by commodity
CommodityCodeProdSum <- aggregate(ProducersValue ~ CommodityCode, data = PCEBridge_Summary_byCom, sum)
colnames(CommodityCodeProdSum) <- c("CommodityCode","ProducersValuebyCommodityCode")

# Add Sums as columns and then find the allocation by dividing each row
PCEBridge_Summary_byCom <- merge(PCEBridge_Summary_byCom, CommodityCodeProdSum)

LineToComCodeMapping <- PCEBridge_Summary_byCom[,c(1,3)]


```


#4. Compare Summary Bridge PCE producer value by commodity vs. col F10 from US Use table 
```{r}
# Add final demand values from Use to line code mapping

bridge_v_F10 <- merge(LineToComCodeMapping, U_F10)
bridge_v_F10 <- merge(bridge_v_F10, PCEBridge_Summary_byCom, by = "CommodityCode" )
# Keep a set of rows where the combintion of commodity code, F10, and ProducersValueByCommodityCode form a unique set
bridge_v_F10 <- unique(bridge_v_F10[c('CommodityCode', 'U_Prod_F10', 'ProducersValuebyCommodityCode')])
bridge_v_F10$U_F10_to_Bridge_Ratio <- bridge_v_F10$U_Prod_F10/bridge_v_F10$ProducersValuebyCommodityCode

outliers_index <- which(abs(bridge_v_F10$U_F10_to_Bridge_Ratio)>1.001 | abs(bridge_v_F10$U_F10_to_Bridge_Ratio) < 0.999)
print("The table below shows any CommodityCodes for which the Summary Bridge data and Use table F10 column have different totals")
print(bridge_v_F10[outliers_index,])


```



#5. Sort PCE Bridge by line and purchaser value to compare with SAPCE3 values
```{r}


# Sort and find ratios by PCE Line Code
PCEBridge_Summary_byLine <- PCEBridge_Summary[order(PCEBridge_Summary$LineCode),]

# Find sum of Purchasers by Line Code
# First convert to numeric
PCEBridge_Summary_byLine$PurchasersValue <- as.numeric(PCEBridge_Summary_byLine$PurchasersValue)
PCEBridge_Summary_byLine[which(is.na(PCEBridge_Summary_byLine$PurchasersValue))] <- 0 # replace possbile NAs from character conversion with 0s

LineCodeSum <- aggregate(PurchasersValue ~ LineCode, data = PCEBridge_Summary_byLine, sum)
colnames(LineCodeSum) <- c("LineCode","PurchasersValuebyLineCodeSum")

# Add ProducersValuebyLineCodeSum as a column and then find the allocation by dividing each row
PCEBridge_Summary_byLine <- merge(PCEBridge_Summary_byLine, LineCodeSum, by = "LineCode")

#PCEBridge_Summary_byLine$Mapping_ratio <- PCEBridge_Summary_byLine$ProducersValue/PCEBridge_Summary_byLine$ProducersValuebyLineCodeSum

```


#6. Compare Summary Bridge Purchaser values vs SAPCE 3 values by line code
```{r}

# Divide SAPCE 3 data by 1e6 to keep units consistent
US_PCE$`2017` <- US_PCE$`2017`/1e6

bridge_v_SAPCE3 <- merge(PCEBridge_Summary_byLine, US_PCE, by = "LineCode")
# Keep relevant columns for comparison
bridge_v_SAPCE3 <- bridge_v_SAPCE3[,c("LineCode","PurchasersValuebyLineCodeSum","2017")]
# Get rid of repeat rows
bridge_v_SAPCE3 <- unique(bridge_v_SAPCE3[,c(1,2,3)])

colnames(bridge_v_SAPCE3) <- c("LineCode","BridgePurValbyLineCodeSum", "US_PCE_PurVal")
# Reorder to keep the same order as the other comparison, i.e., bridge as the denominator of the ratio
bridge_v_SAPCE3 <- bridge_v_SAPCE3[,c(1,3,2)]

bridge_v_SAPCE3$US_PCE_to_Bridge_Ratio <- bridge_v_SAPCE3$US_PCE_PurVal/bridge_v_SAPCE3$BridgePurValbyLineCodeSum


outliers_index <- which(abs(bridge_v_SAPCE3$US_PCE_to_Bridge_Ratio)>1.001 | abs(bridge_v_SAPCE3$US_PCE_to_Bridge_Ratio) < 0.999)
print("The table below shows any LineCodes for which the US PCE data and the Summary Bridge data have different totals")
print(bridge_v_SAPCE3[outliers_index,])


```